---
title: "R Notebook"
Author: "Odalys Barrientos & Fatima Gohar"
output: html_notebook
---

# Lab 5: Get Me Out(lier) of Here!

## Problem 1
```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)

quantile(x, 0.25) # first quartile
quantile(x,0.5) 
quantile(x, 0.75) # third quartile
```

## Problem 2
```{r}
#` This function takes a vector containing a sample
#`
#` @param x the vector containing the sample
#`
#` @return a vector with three components the first quartile, the third quartile, and the interquartile
quartiles<- function (x=vector()){
  firstquart<- quantile(x, 0.25)
  thirdquart<- quantile(x, 0.75)
  interrange <- thirdquart-firstquart
  quarts<- c(firstquart, thirdquart, interrange)
  return(quarts)
}
```

```{r}
quartiles(x)

Q1 <- quantile(x, 0.25)
Q3 <- quantile(x, 0.75)
interrange <- Q3-Q1

Q1
Q2
interrange
```

## Problem 3
According to Tukey's rule the outlier is 28.7. This value does not lie within the range of out.top and out.bottom. 
```{r}
outtop <-Q1-1.5*interrange
outbottom <- Q3+1.5*interrange

outtop
outbottom
```

## Problem 4
```{r}
#` This function takes a vector containing a sample tells which elemnts of the sample are outliers 
#`
#` @param x a vector containing the sample 
#`
#` @return a Boolean vector with TRUE at the elements cooresponding to outlier observations and FALSE everywhere else
tukey.outlier <- function(x=vector()){
  outcomes <- ifelse((x < ((quartiles(x)[1])-1.5*(quartiles(x)[3])) | (x > ((quartiles(x)[2])+1.5*(quartiles(x)[3])))), TRUE, FALSE)
  return(outcomes)
  }
```

```{r}
tukey.outlier(x)

which(tukey.outlier(x))

x[11]
```

## Problem 5
```{r}
#` This function test if your guesses of outliers against tukey's test of outliers
#`
#` @param  the function tukey.outlier which tests for ourliers. 
#`
#` @return TRUE if the function correctly identifies the outliers and FALSE otherwise
test.tukey.outlier <- function(fun = tukey.outlier){
  x<-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  outlier<-fun(x)
  correct <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  is.correct<-sum(correct==outlier)==length(x)
  return(is.correct)
}

test.tukey.outlier(fun=tukey.outlier)
```

## Problem 6
```{r}
z <- -x
```
For -x the value -28.7 is an out lier. 

## Problem 7 
```{r}
#` This function test if your guesses of outliers against tukey's test of outliers
#`
#` @param  the function tukey.outlier 
#`
#` @return TRUE if the function correctly identifies the outliers and FALSE otherwse
test.tukey.outlier <- function(fun = tukey.outlier){
  correct.z <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  correct <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  TorF <- ifelse(correct == fun, TRUE, FALSE)
  TorF.z <- ifelse(correct.z == fun, TRUE, FALSE)
  outcomes <- c(TorF, TorF.z)
  return(outcomes)
}

test.tukey.outlier(tukey.outlier(z))
```

## Problem 8
```{r}
tukey.outlier(z)
which(tukey.outlier(z))
z[11]
```

## Problem 9
For vector y the out lier should be 2.87.
```{r}
y <- 0.1*x
y

outtop.y <-quantile(y,.25)-1.5*(quantile(y,.75)-quantile(y,.25))
outbottom.y <- quantile(y,.75)+1.5*(quantile(y,.75)-quantile(y,.25))

outtop.y
outbottom.y
```
<<<<<<< HEAD

## Problem 10
````{r}
#` This function test if your guesses of outliers against tukey's test of outliers
#`
#` @param  the function tukey.outlier 
#`
#` @return TRUE if the function correctly identifies the outliers and FALSE otherwse
test.tukey.outlier <- function(fun = tukey.outlier){
  correct.y <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  correct.z <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  correct <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  TorF <- ifelse(correct == fun, TRUE, FALSE)
  TorF.z <- ifelse(correct.z == fun, TRUE, FALSE)
  TorF.y <- ifelse(correct.y == fun, TRUE, FALSE)
  outcomes <- c(TorF, TorF.z, TorF.y)
  return(outcomes)
}

test.tukey.outlier(tukey.outlier(y))
```

## Problem 11
```{r}
tukey.outlier(y)
which(tukey.outlier(y))
y[11]
```
=======
For -x the value -28.7 is an outlier. 

## Problem 7
>>>>>>> ec5b20e9880f12b2803be19a5297094b200e1f03

```{r}
#` This function test if your guesses of outliers against tukey's test of outliers
#`
#` @param  the function tukey.outlier which tests for ourliers. 
#`
#` @return TRUE if the function correctly identifies the outliers and FALSE otherwise
test.tukey.outlier <- function(fun = tukey.outlier){
  x<-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
  neg.x<-c(-2.2, -7.8, 4.4, 0.0, 1.2, -3.9, -4.9, 5.7, 7.9, 4.9, -28.7, -4.9)
  outlier<-fun(x)
  negativeoutlier<-fun(neg.x)
  correct <- c(FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,FALSE,TRUE,FALSE)
  is.correct<-sum(correct==outlier)==length(x)
  is.corr<-sum(correct==negativeoutlier)==length(x)
  return(list(is.correct, is.corr))
}

```
## Problem 8

```{r}
test.tukey.outlier(fun=tukey.outlier)
```

## Problem 9

```{r}
x<-c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
0.1*x
```
The outlier is 2.87. 
